---
title: Growth, survival and regeneration of _Calluna vulgaris_ in a common garden
author:
  - id: aut1
    name:
      given: Richard J.
      family: Telford
    orcid: 0000-0001-9826-3076
    email: richard.telford@uib.no
    affiliations:
      - ref: aff1
  - id: aut2
    name:
      given: Vigdis
      family: Vandvik
    orcid: 0000-0003-4651-4798
    email: vigdis.vandvik@uib.no
    affiliations:
      - ref: aff1
  - id: aut3
    name:
      given: Liv Guri
      family: Velle
    orcid: 0000-0003-0433-5337
    email: liv.guri.velle@moreforsking.no
    affiliations:
      - ref: aff2
  - id: aut4
    name:
      given: A. N.
      family: Other
    affiliations: []
date: today
header-includes:
  - \usepackage{lineno}
  - \linenumbers
format:
#  html: default
  pdf: default
documentclass: article
classoption: a4paper
bibliography: extra/calluna2.bib
csl: extra/elsevier-harvard_rjt.csl
execute:
  error: false
  echo: false
affiliations:
  - id: aff1
    name: Department of Biosciences, University of Bergen, Postbox 7803 N-5020, Bergen,
      Norway
  - id: aff2
    name: 'Møreforsking AS Ålesund: Alesund, Norway'
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)
library(gt)
library(targets)
library(plume)
theme_set(theme_bw())
```

```{r}
#| label: load
tar_load(metadata)
```

```{r}
#| label: lat-range
lat_range <- range(metadata$lat) |> diff() |> round(1)
```

# Abstract

*Calluna vulgaris* is widely geographically and environmentally distributed species, keystone species in threatened coastal heathlands.
Economically important.
Within heathlands; geographic variation in growth, response to fire, resprouting, in seed germination, affects productivity and grazing resources.
Management implications for burning rotations.
Seedlings are vulnerable to drought; resprouting and growth rate affects ecosystem resilience to drought.
There is genetic variation but unclear to what extent that is genetic or plastic response to climate, and translates into growth and demographic variation.
To answer this, we did a common garden experiment: 8 populations, `r lat_range`° latitudinal gradient, grown in greenhouse for three years, common garden for 7-8 years, measured survival, growth, fertility, then burnt to assess re-sprouting ability.

# Introduction

Hypothesis: variation is due to populations being locally adapted (vs null hypothesis of plasticity).
Plant growth rate lower in colder-climate populations (chk with litt) More compacts growth forms in the north Earlier sexual recruitment in the north (one mode of reprod. only) Seedling drought etc tolerance higher in the south (i.e., lower mortality) Figures, tables, tests: Site map with heathland belt and Lygra and Tarva inserted (example plot of vegetative and seed recruitment after e.g., 2 yrs) Growth through life (height\~time\*pop rand=mother, or a composite; and also report correlation of the size parameters) Growth form through life (#branches/height) form\~time\*pop rand=mother split inside/out Biomass (simple box plot per pop) Flowering\~size\*pop logistic Mortality\~size\*pop logistic Respouting\~pop --\>

# Methods

*Calluna vulgaris* seeds were collected from eight sites spanning the latitudinal, and climatic, gradient of Norwegian coastal heathlands in MONTH 20XX (@tbl-sites, @fig-site-map).
At each site, ripe seeds were collected from X X plants, spaced at least Xm apart, in X blocks, spaced Xm apart.

FIELD COLLECTION DETAILS

@Brun2022

```{r}
#| label: tbl-sites
#| tbl-cap: Locations _Calluna_ seeds were collect from. GGD5 is growing degrees days above 5°C [@Brun2022; @Brun2022a]
#| include: true
metadata |> 
  select(-site) |> 
  relocate(code) |> 
  arrange(desc(lat)) |> 
  gt() |> 
  cols_label(code = "Code", lat = "Latitude °N", long = "Longitude °E", gdd5 = "GDD5") |> 
  fmt_number(c(lat, long), decimals = 2)
```

```{r}
#| label: fig-site-map
#| fig.cap: The potential distribution of coastal heathands in Norway [ref] with the seed collection sites and the common garden shown. See @tbl-sites for site codes.

tar_read(site_map)
```

GREENHOUSE DETAILS X seeds from each maternal plant were sown.
Plant height and canopy diameter was measured at XXX XXX months.
At XX month, one randomly selected individual from each family was harvested and the above ground biomass measured.

COMMON GARDEN DETAILS After XX years, a random selection of the surviving plants were planted in a common garden.
Plant height was again measured when the plants were XXX years old.

BURNING and RESPROUTING details

In MONTH 20XX the common garden was burned \[DETAILS\].
We identified plants that had re-sprouted in MONTH 20XX.

## Numerical methods

All analyses were done in R v. `r getRversion()` [@R] and used tidyverse v. `r packageVersion("tidyverse")` [@tidyverse] for data manipulation.
Packages targets v. `r packageVersion("targets")` [@targets], quarto `r packageVersion("quarto")` [@quarto] and renv v. `r packageVersion("renv")` [@renv] were used to ensure reproducibility.
Code to reproduce all the analyses is archived at <https://github.com/livguri/Callunachildren>.
Data are archived at XXX

# Results

growth @fig-height-plot

```{r}
#| label: fig-height-plot
#| fig.cap: Height of _Calluna_ plant by site at different time periods.

tar_read(height_plot)
```

hight by mother @fig-height25-mum

```{r}
#| label: fig-height25-mum
#| fig.cap: Height at 2.5 years by mother
tar_read(height2.5_mum_plot)
```

mortality @fig-mortality-plot

```{r}
#| label: fig-mortality-plot
#| fig.cap: Mortality by mother at 2.5 years.
tar_read(mortality_plot)
```

# Discussion

# Author Contributions

```{r}
#| label: author-contributions
#| message: false
#| output: asis
tar_read(plume_authors)$get_contributions(roles_first = TRUE, by_author = TRUE) 
```

## Acknowledgements

landpress

# References
