---
title: "Calluna 6000"
author: "Richard J. Telford"
date: "`r format(Sys.Date(), '%d %B %Y')`"
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output: pdf_document
documentclass: article
classoption: a4paper
bibliography: extra/calluna.bib
csl: extra/elsevier-harvard_rjt.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
theme_set(theme_bw())
```
# Abstract

_Calluna vulgaris_ is widely geographically and environmentally distributed species, keystone species in threatened coastal heathlands. 
Economically important. Within heathlands; geographic variation in growth, response to fire, resprouting, in seed germination, affects productivity and grazing resources. 
Management implications for burning rotations. 
Seedlings are vulnerable to drought; resprouting and growth rate affects ecosystem resilience to drought. 
There is genetic variation but unclear to what extent that is genetic or plastic response to climate, and translates into growth and demographic variation. 
To answer this, we did a common garden experiment: 8 populations, XX degree latitudinal gradient, grown in greenhouse for three years, common garden for 7-8 years, measured survival, growth, fertility, then burnt to assess resprouting ability. 
(resprout as result or as a premise).

## Introduction 

Hypothesis: variation is due to populations being locally adapted (vs null hypothesis of plasticity).
        Plant growth rate lower in colder-climate populations (chk with litt)
        More compacts growth forms in the north
        Earlier sexual recruitment in the north (one mode of reprod. only)
        Seedling drought etc tolerance higher in the south (i.e., lower mortality) 
    Figures, tables, tests:
        Site map with heathland belt and Lygra and Tarva inserted (example plot of vegetative and seed recruitment after e.g., 2 yrs)
        Growth through life (height~time*pop rand=mother, or a composite; and also report correlation of the size parameters)
        Growth form through life (#branches/height) form~time*pop rand=mother split inside/out
        Biomass (simple box plot per pop)
        Flowering~size*pop logistic
        Mortality~size*pop logistic
        Respouting~pop


## Methods

```{r site-table}
readd(meta) %>% 
  arrange(Lat) %>% 
  knitr::kable(caption = "Seed source information", digits = 2)
```



```{r site-map, fig.cap = "Site map"}
readd(site_map)
```


## Results

```{r height-plot, fig.cap = "Height of _Calluna_ plant by site at different time periods."}
readd(height_plot)
```

```{r height2.5-mum, fig.cap = "Height at 2.5 years by mother"}
readd(height2.5_mum)
```


```{r mortality-plot, fig.cap = "Mortality by mother at 2.5 years."}
readd(mortality_plot)
```
