---
title: Calluna 6000
author:
  - id: aut1
    name:
      given: Richard J.
      family: Telford
    orcid: 0000-0001-9826-3076
    email: richard.telford@uib.no
    affiliations:
      - ref: aff1
  - id: aut2
    name:
      given: Vigdis
      family: Vandvik
    orcid: 0000-0003-4651-4798
    email: vigdis.vandvik@uib.no
    affiliations:
      - ref: aff1
  - id: aut3
    name:
      given: Liv Guri
      family: Velle
    orcid: 0000-0003-0433-5337
    email: liv.guri.velle@moreforsking.no
    affiliations:
      - ref: aff2
  - id: aut4
    name:
      given: A. N.
      family: Other
    affiliations: []
date: today
header-includes:
  - \usepackage{lineno}
  - \linenumbers
format: html
documentclass: article
classoption: a4paper
bibliography: extra/calluna.bib
csl: extra/elsevier-harvard_rjt.csl
execute:
  error: false
  echo: true
affiliations:
  - id: aff1
    name: Department of Biosciences, University of Bergen, Postbox 7803 N-5020, Bergen,
      Norway
  - id: aff2
    name: 'Møreforsking AS Ålesund: Alesund, Norway'
---

```{r}
#| label: setup
#| include: false
library(ggplot2)
library(dplyr)
library(targets)
theme_set(theme_bw())
```
# Abstract

_Calluna vulgaris_ is widely geographically and environmentally distributed species, keystone species in threatened coastal heathlands. 
Economically important. Within heathlands; geographic variation in growth, response to fire, resprouting, in seed germination, affects productivity and grazing resources. 
Management implications for burning rotations. 
Seedlings are vulnerable to drought; resprouting and growth rate affects ecosystem resilience to drought. 
There is genetic variation but unclear to what extent that is genetic or plastic response to climate, and translates into growth and demographic variation. 
To answer this, we did a common garden experiment: 8 populations, XX degree latitudinal gradient, grown in greenhouse for three years, common garden for 7-8 years, measured survival, growth, fertility, then burnt to assess resprouting ability. 
(resprout as result or as a premise).

## Introduction 

Hypothesis: variation is due to populations being locally adapted (vs null hypothesis of plasticity).
        Plant growth rate lower in colder-climate populations (chk with litt)
        More compacts growth forms in the north
        Earlier sexual recruitment in the north (one mode of reprod. only)
        Seedling drought etc tolerance higher in the south (i.e., lower mortality) 
    Figures, tables, tests:
        Site map with heathland belt and Lygra and Tarva inserted (example plot of vegetative and seed recruitment after e.g., 2 yrs)
        Growth through life (height~time*pop rand=mother, or a composite; and also report correlation of the size parameters)
        Growth form through life (#branches/height) form~time*pop rand=mother split inside/out
        Biomass (simple box plot per pop)
        Flowering~size*pop logistic
        Mortality~size*pop logistic
        Respouting~pop


## Methods

sites @tbl-sites

```{r}
#| label: tbl-sites
#| tbl-cap: "Seed source information"
tar_read(metadata) |> 
  arrange(Lat) |> 
  knitr::kable(digits = 2)
```

site map @fig-site-map

```{r}
#| label: fig-site-map
#| fig.cap: Site map. See @tbl-sites for site codes.
tar_read(site_map)
```


## Results

growth @fig-height-plot

```{r}
#| label: fig-height-plot
#| fig.cap: Height of _Calluna_ plant by site at different time periods.
tar_read(height_plot)
```

hight by mother @fig-height-25-mum

```{r}
#| label: fig-height25-mum
#| fig.cap: Height at 2.5 years by mother
tar_read(height2.5_mum_plot)
```

mortality @fig-mortality-plot

```{r}
#| label: fig-mortality-plot
#| fig.cap: Mortality by mother at 2.5 years.
tar_read(mortality_plot)
```

# Author Contributions

```{r}
#| label: author-contributions
#| output: asis
library(plume)
authors <- "authors.csv"
aut <- PlumeQuarto$new(read.csv(authors))

aut$get_contributions(role_first = TRUE, name_list = TRUE) |> 
  cat()
```

